{\normalsize Seminar: Formal Specification} \\[1ex] Security of Multithreaded Programs by Compilation\cite{Barthe07}

### 2012-01-30 paper/paper.tex
moved some sections to new pages
fixed spelling
corrected errors found by friends

The idea Barthe, Rezk, Russo and Sabelfeld phrased in \cite{Barthe07} was to close
this leak (formally called \textit{covert channel}) by annotating the byte-code
of the program\footnote{most mobile devices use some sort of byte-code e.g. Android} with security
labels according to the confidentiality of the information, which causes the scheduler to hide
threads that leak information from the attacker.

@@

The idea is to \textit{"track and regulate"}\cite{Zdancewic04} where information flows to
prevent the leak of secret information. The ways on which the information flows
through the system are called channels. Channels that are not indented to transport
information are called covered channels. These covered channels can
be classified into the following categories \cite{Sabelfeld03} (in the following examples $h$ is a secret
variable and $l$ a variable that is public):

@@

The notion of confidentiality that is used in most work in the area of information-flow security is \textit{noninterference}.
This policy is defined in various ways but the essence is, that it requires \textit{"that secret information
[do] not affect [the] publicly observable behavior of a system"}\cite{Zdancewic04}\footnote{Even if the attacker has full access to the
source code of the program.}. For most real world
applications this policy is far too strict, because it forbids useful programs like password checkers.\footnote{A
password checker needs to reveal whether the user input was correct or not.} Approaches that allow
controlled release of information are called declassification. The kinds of declassification are \textit{what}
information is released, \textit{who} released this information, \textit{where} in the system the information
is released and \textit{when} the information is released (c.f. \cite{Sabelfeld05}). For these approaches it is necessary that an
active attacker can only know as much as a passive attacker.

@@

\subsection{Syntax and Semantics of multithreaded programs}
\label{sec:syntaxsemantics}
A program is viewed as an abstract thing, which consists of a set of
program points $\mathcal{P}$ with a distinguished entry ($1$) and exit
(\texttt{exit}) point and a function $P$ that maps program points to
instructions.

@@

At this point a first simplification can be made. Looking a state $s \in \text{ConcState}$ we can
first extract the active threads (\texttt{s.act}) by taking the domain of the first
component. According to a security environment we can classify these threads,
with respect to their current program point (\texttt{s.pc(tid)} where $tid \in \text{Thread}$),
into \textit{low} threads if the current program point is low. In \textit{high}
threads iff the current program point is high. In \textit{always high} threads
if the current program point is always high and in \textit{hidden} threads, if
the current program point is high but not always high.

The latter two are the interesting ones. If a thread is \textit{always high} it
can not leak any information into low, because it never gets in touch
with low program points. So these threads can safely be interleaved between all
other threads by the scheduler.

The \textit{hidden} threads are the ones we have to care about. These contain
obviously high information in the current program point, but have afterwards
instructions that deal with low information. Since the attacker can watch the
low part of the memory, chances are good that he can deduce high information
through looking at changing low outputs. To prevent indirect flows that are introduced by
these hidden threads, the scheduler will be modified to treat these threads
in a special way. This will be done by "hiding" these threads, therefore comes
the name of them.

@@

\begin{enumerate}
\item It always picks an active thread
\item If there is a hidden thread, always choose high or always high threads\footnote{With this constraint the interleaving of always high threads is realized}
\item Only use low information and the low part of the history to choose a
      new thread
\end{enumerate}

@@

\begin{align*}
\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] = \mathtt{start} pc \\
fresht_{se(i)}(s) = ntid \\ s(ctid).[pc := i + 1] = \sigma'}
{s,h \leadsto_{conc} s.[lst(ctid) := \sigma', lst(ntid) := \lambda_{init}(pc)],
\langle ctid, se(i)\rangle :: h}
\end{align*}
\caption{Multithreaded execution.}
\label{fig:multithreaded-execution}
\end{figure}

The scheduler is allowed to pick a new thread after every transition of $\leadsto_{seq}$. The
intuition of the first rule is, that the scheduler picks a new thread $coed$ which is at
program point $i$. This program point maps to an sequential instruction. The execution of
this instruction leads to a local state $\sigma$ and a global memory $\mu$. If this is the case
and the instruction $i$ was not the last instruction (i.e. $\sigma.pc \neq \mathtt{exit}$),
then the concurrent transition can be made. In this transition the concurrent state is updated
according to the results of the sequential execution and the thread id furthermore security level of $i$
is recorded in the history.

The second rule covers the case in which the thread picked by the scheduler has only the current
instruction before terminating (i.e. $\sigma.pc = \mathtt{exit}$). In the concurrent transition
the current thread id is removed from the concurrent state and everything else is like in the
first rule.

The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
\footnote{It is assumed that $fresh_l(tid) \neq fresh_{l'}(tid) \Leftrightarrow l \neq l'$.}
takes a set of thread identifiers and returns a new thread identifier at level $l$ and $\lambda_{init} : \mathcal{P}
\rightarrow \text{LocState}$ produces an initial state with the program pointer at the given program point.
If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
identifier is picked w.r.t. the security level of $i$, then program counter of the current thread
is increased by one, to step to the next instruction. The resulting concurrent state includes the
updated state of the current thread and the initial state for the new thread. The global memory
is not modified and the history is extended with the current thread id and the security level of $i$.

@@

This type system is extended by the following rules, to support multithreading:

@@

A program is typeable in this type system (written $\mathcal{S}, se \vdash P$),
where $\mathcal{S}$ is a function $\mathcal{S}: \mathcal{P} \rightarrow \text{LType}$
that maps a local type to every program point
and $se$ a security environment, iff $\mathcal{S}$ maps every initial program point
\footnote{including the ones of spawned threads} to $\textsc{t}_{init}$ and that for every program point $j$, which is an
successor of $i$, there is a type $s \in \text{LType}$, such that $\textsc{s}$ is lower bounded
by $\mathcal{S}(j)$ and $se, i \vdash \mathcal{S}(i) \Rightarrow \textsc{s}$ holds.

@@

\subsection{Instantiation}
\label{sec:instantiation}
To demonstrate how the framework can be used, it was instantiated with a simple
assembly language, which is given by the follow grammar:

@@

According to this labels we can define the dependence regions for the source language:

@@

\begin{definition}
For every branching point $c$ in the source program such that $\vdash_\circ [c]_\bullet^n$, the next function
is defined as $\forall k \in tregion(n) .\ next(k) = jun(n)$.
\end{definition}

@@

\section{Comparison with other approaches/further work}
\label{sec:furtherwork}
In this section I will first give a short overview of two other approaches and then
compare them with the approach presented before and draw a conclusion.

\subsection{Observational Determinism for Concurrent Program Security\cite{Zdancewic03}}
Zdancewic and Myers approach has to goal to provide a secure concurrent language
that has a general and realistic support for concurrency and whose security can
be checked statically.

@@

In the end Barthe et. al. presented a framework with reaches many goals (like permissiveness, statically check-able
and requires no intervention from the programmer) but needs to modify the scheduler interface, which makes it
hard to use for real programs. In addition they did not how to implement the type-annotations in a real
language.

\bibliography{bibliography}

>>>>

The idea Barthe, Rezk, Russo and Sabelfeld phrased in \cite{Barthe07} was to close
this leak (formally called \textit{covert channel}) by annotating the byte-code
of the program\footnote{Most mobile devices use some sort of byte-code e.g. Android.} with security
labels according to the confidentiality of the information, which causes the scheduler to hide
threads that leak information from the attacker.

@@

The idea is to \textit{"track and regulate"}\cite{Zdancewic04} where information flows to
prevent the leak of secret information. The ways on which the information flows
through the system are called \textit{channels}. Channels that are not intended to transport
information are called \textit{covered channels}. These covered channels can
be classified into the following categories \cite{Sabelfeld03} (in the following examples $h$ is a secret
variable and $l$ a variable that is public):

@@

The notion of confidentiality that is used in most work in the area of in\-for\-ma\-tion-flow security is \textit{noninterference}.
This policy is defined in various ways but the essence is, that it requires \textit{"that secret information
[do] not affect [the] publicly observable behavior of a system"}\cite{Zdancewic04}\footnote{Even if the attacker has full access to the
source code of the program.}. For most real world
applications this policy is far too strict, because it forbids useful programs like password checkers.\footnote{A
password checker needs to reveal whether the user input was correct or not.} Approaches that allow
controlled release of information are called declassification. The kinds of declassification are \textit{what}
information is released, \textit{who} released this information, \textit{where} in the system the information
is released and \textit{when} the information is released (c.f. \cite{Sabelfeld05}). For these approaches it is necessary that an
active attacker can only know as much as a passive attacker.

@@

\subsection{Syntax and Semantics of multithreaded programs}
\label{sec:syntaxsemantics}
A program is viewed as an abstract thing, which consists of a set of
program points $\mathcal{P}$ with a distinguished entry ($1$) and exit
point (\texttt{exit}) and a function $P$ that maps program points to
instructions.

@@

At this point a first simplification can be made. Looking a state $s \in \text{ConcState}$ we can
first extract the active threads (\texttt{s.act}) by taking the domain of the first
component. According to a security environment we can classify these threads,
with respect to their current program point (\texttt{s.pc(tid)} where $tid \in \text{Thread}$),
into \textit{low} threads if the current program point is low. In \textit{high}
threads if the current program point is high. In \textit{always high} threads
if the current program point is always high and in \textit{hidden} threads, if
the current program point is high but not always high.

The last two are the interesting ones. If a thread is \textit{always high} it
can not leak any information into low, because it never gets in touch
with low program points. So these threads can safely be interleaved between all
other threads by the scheduler.

The \textit{hidden} threads are the ones we have to care about. These contain
obviously high information in the current program point, but have subseqent
instructions that deal with low information. Since the attacker can watch the
low part of the memory, chances are good that he can deduce high information
through looking at changing low outputs. To prevent indirect flows that are introduced by
these hidden threads, the scheduler will be modified to treat these threads
in a special way. This will be done by "hiding" these threads, therefore comes
the name of them.

@@

\begin{enumerate}
\item It always picks an active thread
\item If there is a hidden thread, always choose high or always high threads\footnote{With this constraint the interleaving of always high threads is realized.}
\item Only use low information and the low part of the history to choose a
      new thread
\end{enumerate}

@@

\begin{align*}
\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] = \mathtt{start}\ pc \\
fresht_{se(i)}(s) = ntid \\ s(ctid).[pc := i + 1] = \sigma'}
{s,h \leadsto_{conc} s.[lst(ctid) := \sigma', lst(ntid) := \lambda_{init}(pc)],
\langle ctid, se(i)\rangle :: h}
\end{align*}
\caption{Multithreaded execution.}
\label{fig:multithreaded-execution}
\end{figure}

The scheduler is allowed to pick a new thread after every transition of $\leadsto_{seq}$. The
intuition of the first rule is, that the scheduler picks a new thread $coed$ which is at
program point $i$. This program point maps to an sequential instruction. The execution of
this instruction leads to a local state $\sigma$ and a global memory $\mu$. If this is the case
and the instruction $i$ was not the last instruction (i.e. $\sigma.pc \neq \mathtt{exit}$),
then the concurrent transition can be made. In this transition the concurrent state is updated
according to the results of the sequential execution and the thread identifier, furthermore security level of $i$
is recorded in the history.

The second rule covers the case in which the thread picked by the scheduler has only the current
instruction before terminating (i.e. $\sigma.pc = \mathtt{exit}$). In the concurrent transition
the current thread identifier is removed from the concurrent state and everything else is like in the
first rule.

The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
\footnote{It is assumed that $fresh_l(tid) \neq fresh_{l'}(tid) \Leftrightarrow l \neq l'$.}
takes a set of thread identifiers and returns a new thread identifier at level $l$ and $\lambda_{init} : \mathcal{P}
\rightarrow \text{LocState}$ produces an initial state with the program pointer at the given program point.
If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
identifier is picked w.r.t. the security level of $i$, then program counter of the current thread
is increased by one, to step to the next instruction. The resulting concurrent state includes the
updated state of the current thread and the initial state for the new thread. The global memory
is not modified and the history is extended with the current thread identifier and the security level of $i$.

@@

This type system is extended by the following rules to support multithreading:

@@

A program is typeable in this type system (written $\mathcal{S}, se \vdash P$),
where $\mathcal{S}$ is a function $\mathcal{S}: \mathcal{P} \rightarrow \text{LType}$
that maps a local type to every program point
and $se$ a security environment, iff $\mathcal{S}$ maps every initial program point\footnote{
Including the ones of spawned threads.} to $\textsc{t}_{init}$ and that for every program point $j$, which is an
successor of $i$, there is a type $s \in \text{LType}$, such that $\textsc{s}$ is lower bounded
by $\mathcal{S}(j)$ and $se, i \vdash \mathcal{S}(i) \Rightarrow \textsc{s}$ holds.

@@

\newpage
\subsection{Instantiation}
\label{sec:instantiation}
To demonstrate how the framework can be used, it was instantiated with a simple
assembly language, which is given by the following grammar:

@@

According to this labels we can define the control dependence regions for the source language:

@@

\begin{definition}
For every branching point $c$ in the source program such that $\vdash_\circ~[c]_\bullet^n$, the next function
is defined as $\forall k \in tregion(n) .\ next(k) = jun(n)$.
\end{definition}

@@

\newpage
\section{Comparison with other approaches/further work}
\label{sec:furtherwork}
In this section I will first give a short overview of two other approaches and then
compare them with the approach presented before and draw a conclusion.

\subsection{Observational Determinism for Concurrent Program Security\cite{Zdancewic03}}
Zdancewic and Myers approach has the goal to provide a secure concurrent language
that has a general and realistic support for concurrency and whose security can
be checked statically.

@@

In the end Barthe et. al. presented a framework with reaches many goals (like permissiveness, statically check-able
and requires no intervention from the programmer) but needs to modify the scheduler interface, which makes it
hard to use for real programs. In addition they did not showed how to implement the type-annotations in a real
language.

\newpage
\bibliography{bibliography}
